---
output: 
  beamer_presentation: 
    colortheme: beaver
    fonttheme: structurebold
    incremental: no
    theme: boxes
---

## Introduction:

**Bee local: A comparison of productivity and pathogen load in local vs. Claifornia re-queened colonies**

\bigskip
\bigskip
\bigskip


Andre Burnham, Fiona McLaughlin, **P. Alexander Burnham** & Herman Lehman

 \begin{picture}(200,200)
 \put(1,70){\includegraphics[height=3cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture6.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(155, 288){\includegraphics[height=2.2cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture7.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 

## Acknowledgments

**Co-Authors:**

  * Andre Burnham   
  * Fiona McLaughlin   
  * Dr. Herman Lehman   

**Thank you to:**

  * The Casstevens Family   
  * Nancy Thompson   
  * Samantha Alger   


 \begin{picture}(200,200)
 \put(155,100){\includegraphics[height=3cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture6.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(155, 400){\includegraphics[height=2cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture7.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 
 
   \begin{picture}(200,200)
 \put(155, 700){\includegraphics[height=1.3cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture14.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 

## The Bee Team


 \begin{picture}(200,200)
 \put(-25,60){\includegraphics[height=5cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture5.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(182, 166){\includegraphics[height=5.3cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture20.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 
## Honey Bees are Important

* 30% of the world’s food is derived from pollination (Aizen et al., 2009)   
* Pollinators are responsible for between $235-577 billion (Gallai et al., 2009)   
* Honeybees are responsible for $14 Billion in the USA (Morse & Calderone, 2000)   

 \begin{picture}(200,200)
 \put(60, 70){\includegraphics[height=4.5cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture21.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}




## Threats to Bees:

 \begin{picture}(200,200)
 \put(10, 1){\includegraphics[height=17cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture19.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 

## Honey Bee Pathogens

**VIRUSES:**

* Deformed Wing   
* Black Queen Cell   
* Israeli Acute Paralysis   

**PARASITES:**

* Nosema (ceranae/apis)   
* Varroa Mite   


 \begin{picture}(200,200)
 \put(130, 90){\includegraphics[height=8cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture22.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 
 
## Troubles for Beekeepers (re-queening)

 \begin{picture}(200,200)
 \put(-25,60){\includegraphics[height=5cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture25.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(130, 166){\includegraphics[height=5.3cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture26.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}

## The basic premises behind this study

* Imported VS Local   
* Local Adapation   

 \begin{picture}(200,200)
 \put(-22,50){\includegraphics[height=4cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture3.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(176, 256){\includegraphics[height=4cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture4.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 
 
 ## The basic premises behind this study
 
* Mass-Produced VS Handmade
* Selection by the Breeder   

 \begin{picture}(200,200)
 \put(-23,50){\includegraphics[height=4cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture1.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(158, 256){\includegraphics[height=4cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture2.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 

 

## The question:

*“Are locally-bred queens more successful than imported queens?”*

## Experiemental Design

* 20 colonies re-queened with Californian-bred queens   

* 20 colonies re-queened local-bred (Vermont) queens   

* 2 sites, 10 Local and 10 California for each   

* Sampled for pathogens and productivity measures   

* Sampled at different time points for 3 months  


## Pictures of the Yards

 \begin{picture}(200,200)
 \put(-22,0){\includegraphics[height=6cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture8.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(202, 206){\includegraphics[height=6cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture9.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 

## What we sampled

* Growth:   
    - Colony Mass   
    - Brood Production    
* Foraging:   
    - Pollen Production   
* Pathogens:   
    - Varroa   
    - Nosema spp.   
    - RNA Viruses   







## Our Predictions

* Local queens (colonies) will have higher growth through the season   

* Local queens will be better foragers   

* Local queens (colonies) will have lower pathogen loads   



## Colony Mass (growth)

```{r, echo=FALSE, warning=FALSE}


# Set Working Directory: 
setwd("~/AndreCollaborations/QueenExperimentBurnham")

# Read in Data:
QueenDF <- read.table("2016QueensHam.csv", header=TRUE, sep = ",", stringsAsFactors = FALSE) 



# These dataframes are seperate becuase they do not follow that sampling 
# pattern that the other four varaibles in the previous data frame do (i.e.5
# time steps):
PollenDF <- read.table("PollenQueens.csv", header=TRUE, sep = ",") 
VirusDF <- read.table("RNAVirus.csv", header=TRUE, sep = ",") 

# load plyr for data manipulation and ggplot plotting package and other 
# related packages:
library(plyr)
library(ggplot2)
library(grid)
library(scales)


# summary stats for plotting purposes:
MassSummary <- ddply(QueenDF, c("Origin", "MassDay"), summarise, 
                     n = length(Mass),
                     mean = mean(Mass, na.rm = TRUE),
                     sd = sd(Mass, na.rm = TRUE),
                     se = sd / sqrt(n))

# remove time steps with missing data:
MassSummary <- MassSummary[-c(5,10),]

#Create plot in ggplot for brood 
plot <- ggplot(data = MassSummary, 
                aes(x = MassDay, 
                    y = mean, 
                    group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = "Time (days after requeening)", y = "Colony Mass (lbs)") + coord_cartesian(ylim = c(20,70), xlim = c(5,70)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_classic(base_size = 17) + theme(legend.position=c(0.1, 0.85), panel.border = element_blank(), axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'), axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.y=element_text(margin=margin(0,20,0,0))) + scale_x_continuous(breaks=pretty_breaks(n=10)) 

plot
```


## Frames of Brood (growth)

```{r, echo=FALSE, warning=FALSE}
#------------------------------------------------------------------------
# BROOD

# summary stats for plotting purposes:
BroodSummary <- ddply(QueenDF, c("Origin", "BroodDay"), summarise, 
                     n = length(Brood),
                     mean = mean(Brood, na.rm = TRUE),
                     sd = sd(Brood, na.rm = TRUE),
                     se = sd / sqrt(n))

#------------------------------------------------------------------------

#Create plot in ggplot for brood 
plot1 <- ggplot(data = BroodSummary, 
               aes(x = BroodDay, 
                   y = mean, 
                   group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = "Time (days after requeening)", y = "Frames of Brood") + coord_cartesian(ylim = c(2, 7), xlim = c(-10,70)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_classic(base_size = 17) + theme(legend.position=c(0.1, 0.85), panel.border = element_blank(), axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.y=element_text(margin=margin(0,20,0,0))) + scale_x_continuous(breaks=pretty_breaks(n=10)) 

plot1
```


## Pollen Collection (foraging)

```{r, echo=FALSE, warning=FALSE}
# summary stats for plotting purposes:
VarroaSummary <- ddply(QueenDF, c("Origin", "VarroaDay"), summarise, 
                     n = length(Varroa),
                     mean = mean(Varroa, na.rm = TRUE),
                     sd = sd(Varroa, na.rm = TRUE),
                     se = sd / sqrt(n))

# remove time steps with missing data:
VarroaSummary <- VarroaSummary[-c(5,10),]

#------------------------------------------------------------------------

ggplot(data = VarroaSummary, 
               aes(x = VarroaDay, 
                   y = mean, 
                   group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = "Time (days after requeening)", y = "Varroa Load (mites/300 bees)") + coord_cartesian(ylim = c(0, 12), xlim = c(10,70)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_classic(base_size = 17) + theme(legend.position=c(.85, .85), panel.border = element_blank(), axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_x_continuous(breaks=pretty_breaks(n=5)) + labs(linetype="Queen Origin") + annotate("segment", x = 7.1, xend = 72, y = 3, yend = 3, col="red", lty = "dashed")


```


## Varroa Load (pathogens)

```{r, echo=FALSE, warning=FALSE}
# summary stats for plotting purposes:
NosemaSummary <- ddply(QueenDF, c("Origin", "NosemaDay"), summarise, 
                     n = length(Nosema),
                     mean = mean(Nosema, na.rm = TRUE),
                     sd = sd(Nosema, na.rm = TRUE),
                     se = sd / sqrt(n))

# remove time steps with missing data:
NosemaSummary <- NosemaSummary[-c(4,8),]

#------------------------------------------------------------------------

ggplot(data = NosemaSummary, 
               aes(x = NosemaDay, 
                   y = mean, 
                   group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = "Time (days after requeening)", y = "Nosema Load (spores/bee)") + coord_cartesian(ylim = c(0, 4300000), xlim = c(10,70)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_classic(base_size = 17) + theme(legend.position=c(.15, .85), panel.border = element_blank(), axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_x_continuous(breaks=pretty_breaks(n=5)) + labs(linetype="Queen Origin") + annotate("segment", x = 7.1, xend = 72, y = 1000000, yend = 1000000, col="red", lty = "dashed")
```


## Nosema Load (pathogens)

```{r, echo=FALSE, warning=FALSE}
# summary stats for plotting purposes:
PollenSummary <- ddply(PollenDF, c("Origin", "Week"), summarise, 
                     n = length(Pollen),
                     mean = mean(Pollen, na.rm = TRUE),
                     sd = sd(Pollen, na.rm = TRUE),
                     se = sd / sqrt(n))

#------------------------------------------------------------------------

ggplot(data = PollenSummary, 
               aes(x = Week, 
                   y = mean, 
                   group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = "Time (weeks after requeening)", y = "Pollen Mass (grams)") + coord_cartesian(ylim = c(0, 40), xlim = c(1,9)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.1)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_classic(base_size = 17) + theme(legend.position=c(.2, .85), panel.border = element_blank(), axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_x_continuous(breaks=pretty_breaks(n=9)) + labs(linetype="Queen Origin")

```


## Viral Load (pathogens)

```{r, echo=FALSE, warning=FALSE}
VirusDF$IntensityMult <- VirusDF$Band_Intensity*10^6
VirusDFsplit <- split(VirusDF, VirusDF$Virus)

# summary stats for plotting purposes:
VirusSummary <- ddply(VirusDF, c("Origin", "Virus", "Time"), summarise, 
                     n = length(IntensityMult),
                     mean = mean(IntensityMult, na.rm = TRUE),
                     sd = sd(IntensityMult, na.rm = TRUE),
                     se = sd / sqrt(n))

# split data frame by virus type:
splitVir <- split(VirusSummary, VirusSummary$Virus)


#------------------------------------------------------------------------
#DWV

colors <- c("white", "gray32")

p1 <- ggplot(splitVir$DWV, aes(x=Time, y=mean, fill=Origin)) + 
  geom_bar(stat="identity", col="black",  
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=.4,
                position=position_dodge(.9)) + labs(x="DWV", y = "Viral Load (band intensity)") + theme_classic(base_size = 17) + coord_cartesian(ylim = c(0, 4000000)) + scale_fill_manual(values=colors, name="Queen Origin", labels=c("California", "Local")) + theme(legend.position=c(.5, .9), panel.border = element_blank(),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 

#------------------------------------------------------------------------
#BQCV

p2 <- ggplot(splitVir$BQCV, aes(x=Time, y=mean, fill=Origin)) + 
  geom_bar(stat="identity", col="black",  
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=.4,
                position=position_dodge(.9)) + labs(x="BQCV", y = NULL) + theme_classic(base_size = 17) + coord_cartesian(ylim = c(0, 4000000)) + scale_fill_manual(values=colors, name="Queen Origin", labels=c("California", "Local")) + theme(legend.position=c(3, 3), axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),  panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line.y = element_blank()) +  annotate("segment", x = .5, xend = 2.5, y = 3900000, yend = 3900000) + annotate("segment", x = .5, xend = 1.4, y = 1500000, yend = 1500000) + annotate("segment", x = 1.5, xend = 2.5, y = 3600000, yend = 3600000) + annotate(geom = "text", x = 1.5, y = 4000000, label = "***",cex = 6) + annotate(geom = "text", x = 1, y = 1650000, label = "ns",cex = 5) + annotate(geom = "text", x = 2, y = 3750000, label = "ns",cex = 5)

#------------------------------------------------------------------------
#IAPV

p3 <- ggplot(splitVir$IAPV, aes(x=Time, y=mean, fill=Origin)) + 
  geom_bar(stat="identity", col="black",  
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), 
                width=.4,
                position=position_dodge(.9)) + labs(x="IAPV", y = NULL) + theme_classic(base_size = 17) + coord_cartesian(ylim = c(0, 4000000)) + scale_fill_manual(values=colors, name="Queen Origin", labels=c("California", "Local")) + theme(legend.position=c(3, 3),axis.text.y=element_blank(),
        axis.ticks.y=element_blank(), axis.line.y = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + annotate("segment", x = .5, xend = 2.5, y = 3100000, yend = 3100000) + annotate("segment", x = .5, xend = 1.4, y = 2200000, yend = 2200000) + annotate("segment", x = 1.5, xend = 2.5, y = 2800000, yend = 2800000) + annotate(geom = "text", x = 1.5, y = 3250000, label = "ns",cex = 5) + annotate(geom = "text", x = 1, y = 2300000, label = "**",cex = 6) + annotate(geom = "text", x = 2, y = 2900000, label = "**",cex = 6)

# merge figures into 1 side by side figure:
grid.newpage()
grid.draw(cbind(ggplotGrob(p1), ggplotGrob(p2), ggplotGrob(p3), size = "last"))
```


## In Summary 

* Colony Mass 		= **Higher in Local**   
* Amount of Brood 	= **Higher in Local**   
* Pollen Collection 	= **No Difference**   
* Varroa Load 		= **No Difference**   
* Nosema Load		= **Lower in Local**   
* RNA Viruses 		= **Mixed Results**   


## In Summary

* Overall, colonies re-queened with locally raised queens had higher growth   

* Pollen collection did not seem to be involved in this growth   

* Some pathogens seemed to have less of an effect on local honeybees and others had similar effects across both groups   


## Implications

* Locally-raised queens outperform mass-produced, California queens in their northern environment.   

* This could be evidence for the importance of care in breeding stocks (mass produced vs handmade)   

* **And/Or** This could be evidence for local (genetic) adaptation (imported vs. local)   



## Thank you!

 \begin{picture}(200,200)
 \put(1, 1){\includegraphics[height=30cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture13.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}

## Questions?

 \begin{picture}(200,200)
 \put(1, 1){\includegraphics[height=30cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture13.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 


