---
output: 
  beamer_presentation: 
    colortheme: beaver
    fonttheme: structurebold
    incremental: no
    theme: boxes
---

## Introduction:

**Bee local: A comparison of productivity and pathogen load in local vs. Claifornia re-queened colonies**

\bigskip
\bigskip
\bigskip


Andre Burnham, Fiona McLaughlin, **P. Alexander Burnham** & Herman Lehman

 \begin{picture}(200,200)
 \put(1,70){\includegraphics[height=3cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture6.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(155, 288){\includegraphics[height=2.2cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture7.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 

## Acknowledgments

**Co-Authors:**

  * Andre Burnham   
  * Fiona McLaughlin   
  * Dr. Herman Lehman   

**Thank you to:**

  * The Casstevens Family   
  * Nancy Thompson   
  * Samantha Alger   


 \begin{picture}(200,200)
 \put(155,100){\includegraphics[height=3cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture6.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(155, 400){\includegraphics[height=2cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture7.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 
 
   \begin{picture}(200,200)
 \put(155, 700){\includegraphics[height=1.3cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture14.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 

## The Bee Team


 \begin{picture}(200,200)
 \put(-25,60){\includegraphics[height=5cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture5.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(182, 166){\includegraphics[height=5.3cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture20.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 
## Honey Bees are Important

* 30% of the world’s food is derived from pollination (Aizen et al., 2009)   
* Pollinators are responsible for between $235-577 billion (Gallai et al., 2009)   
* Honeybees are responsible for $14 Billion in the USA (Morse & Calderone, 2000)   

 \begin{picture}(200,200)
 \put(60, 70){\includegraphics[height=4.5cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture21.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}




## Honey Bee Pathogens

**VIRUSES:**

* Deformed Wing   
* Black Queen Cell   
* Israeli Acute Paralysis   

**PARASITES:**

* Nosema (ceranae/apis)   
* Varroa Mite   


 \begin{picture}(200,200)
 \put(130, 90){\includegraphics[height=8cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture22.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 
 
## Troubles for Beekeepers (re-queening)

 \begin{picture}(200,200)
 \put(-25,60){\includegraphics[height=5cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture25.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(130, 166){\includegraphics[height=5.3cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture26.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}

## The basic premises behind this study

* Imported VS Local   
* Local Adapation   

 \begin{picture}(200,200)
 \put(-22,50){\includegraphics[height=4cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture3.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(176, 256){\includegraphics[height=4cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture4.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 
 
 ## The basic premises behind this study
 
* Mass-Produced VS Handmade
* Selection by the Breeder   

 \begin{picture}(200,200)
 \put(-23,50){\includegraphics[height=4cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture1.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(158, 256){\includegraphics[height=4cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture2.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 

 

## The question:

*“Are locally-bred queens more successful than imported queens?”*


## Our Predictions

* Local queens (colonies) will have higher growth through the season   

* Local queens will be better foragers   

* Local queens (colonies) will have lower pathogen loads   



## Experiemental Design

* 20 colonies re-queened with Californian-bred queens   

* 20 colonies re-queened local-bred (Vermont) queens   

* 2 sites, 10 Local and 10 California for each   

* Sampled for pathogens and productivity measures   

* Sampled at different time points for 3 months  


## Pictures of the Yards

 \begin{picture}(200,200)
 \put(-22,0){\includegraphics[height=6cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture8.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}


 \begin{picture}(200,200)
 \put(202, 206){\includegraphics[height=6cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture9.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 

## What we sampled

* Growth:   
    - Colony Mass   
    - Brood Production    
* Foraging:   
    - Pollen Production   
* Pathogens:   
    - Varroa   
    - Nosema spp.   
    - RNA Viruses   

## Data Analysis

```
aov.out <- aov(Nosema ~ Origin * Time + Error(FieldID), 
      data=QueenDF)

summary(aov.out)
```

## Repeated Measures ANOVA output

```
Error: FieldID
            Df    Sum Sq   Mean Sq F value  Pr(>F)   
Origin       1 3.156e+13 3.156e+13   7.972 0.00779 **
Time         1 2.589e+12 2.589e+12   0.654 0.42413   
Origin:Time  1 9.223e+11 9.223e+11   0.233 0.63234   
Residuals   35 1.386e+14 3.959e+12                   
---

Error: Within
            Df    Sum Sq   Mean Sq F value   Pr(>F)    
Time         1 5.620e+10 5.620e+10   0.017 0.896098    
Origin:Time  1 4.934e+13 4.934e+13  15.111 0.000275 ***
---                    

```

## Colony Mass (growth)

```{r, echo=FALSE, warning=FALSE}


# Set Working Directory: 
setwd("~/AndreCollaborations/QueenExperimentBurnham")

# Read in Data:
QueenDF <- read.table("2016QueensHam.csv", header=TRUE, sep = ",", stringsAsFactors = FALSE) 



# These dataframes are seperate becuase they do not follow that sampling 
# pattern that the other four varaibles in the previous data frame do (i.e.5
# time steps):
PollenDF <- read.table("PollenQueens.csv", header=TRUE, sep = ",") 
VirusDF <- read.table("RNAVirus.csv", header=TRUE, sep = ",") 

# load plyr for data manipulation and ggplot plotting package and other 
# related packages:
library(plyr)
library(ggplot2)
library(grid)
library(scales)


# summary stats for plotting purposes:
MassSummary <- ddply(QueenDF, c("Origin", "MassDay"), summarise, 
                     n = length(Mass),
                     mean = mean(Mass, na.rm = TRUE),
                     sd = sd(Mass, na.rm = TRUE),
                     se = sd / sqrt(n))

# remove time steps with missing data:
MassSummary <- MassSummary[-c(5,10),]

#Create plot in ggplot for brood 
plot <- ggplot(data = MassSummary, 
                aes(x = MassDay, 
                    y = mean, 
                    group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = "Time (days after requeening)", y = "Colony Mass (lbs)") + coord_cartesian(ylim = c(20,70), xlim = c(5,70)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_classic(base_size = 17) + theme(legend.position=c(0.1, 0.85), panel.border = element_blank(), axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'), axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.y=element_text(margin=margin(0,20,0,0))) + scale_x_continuous(breaks=pretty_breaks(n=10)) 

plot
```


## Frames of Brood (growth)

```{r, echo=FALSE, warning=FALSE}
#------------------------------------------------------------------------
# BROOD

# summary stats for plotting purposes:
BroodSummary <- ddply(QueenDF, c("Origin", "BroodDay"), summarise, 
                     n = length(Brood),
                     mean = mean(Brood, na.rm = TRUE),
                     sd = sd(Brood, na.rm = TRUE),
                     se = sd / sqrt(n))

#------------------------------------------------------------------------

#Create plot in ggplot for brood 
plot1 <- ggplot(data = BroodSummary, 
               aes(x = BroodDay, 
                   y = mean, 
                   group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = "Time (days after requeening)", y = "Frames of Brood") + coord_cartesian(ylim = c(2, 7), xlim = c(-10,70)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_classic(base_size = 17) + theme(legend.position=c(0.1, 0.85), panel.border = element_blank(), axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.y=element_text(margin=margin(0,20,0,0))) + scale_x_continuous(breaks=pretty_breaks(n=10)) 

plot1
```




## Varroa Load (pathogens)

```{r, echo=FALSE, warning=FALSE}
# summary stats for plotting purposes:
NosemaSummary <- ddply(QueenDF, c("Origin", "NosemaDay"), summarise, 
                     n = length(Nosema),
                     mean = mean(Nosema, na.rm = TRUE),
                     sd = sd(Nosema, na.rm = TRUE),
                     se = sd / sqrt(n))

# remove time steps with missing data:
NosemaSummary <- NosemaSummary[-c(4,8),]

#------------------------------------------------------------------------

ggplot(data = NosemaSummary, 
               aes(x = NosemaDay, 
                   y = mean, 
                   group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = "Time (days after requeening)", y = "Nosema Load (spores/bee)") + coord_cartesian(ylim = c(0, 4300000), xlim = c(10,70)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.9)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_classic(base_size = 17) + theme(legend.position=c(.15, .85), panel.border = element_blank(), axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_x_continuous(breaks=pretty_breaks(n=5)) + labs(linetype="Queen Origin") + annotate("segment", x = 7.1, xend = 72, y = 1000000, yend = 1000000, col="red", lty = "dashed")
```


## Nosema Load (pathogens)

```{r, echo=FALSE, warning=FALSE}
# summary stats for plotting purposes:
PollenSummary <- ddply(PollenDF, c("Origin", "Week"), summarise, 
                     n = length(Pollen),
                     mean = mean(Pollen, na.rm = TRUE),
                     sd = sd(Pollen, na.rm = TRUE),
                     se = sd / sqrt(n))

#------------------------------------------------------------------------

ggplot(data = PollenSummary, 
               aes(x = Week, 
                   y = mean, 
                   group = Origin)
) + geom_point(size=3) + scale_colour_manual(values = c("black", "black")) + labs(x = "Time (weeks after requeening)", y = "Pollen Mass (grams)") + coord_cartesian(ylim = c(0, 40), xlim = c(1,9)) + geom_errorbar(aes(ymin = mean - se, ymax = mean + se, width = 0.1)) + geom_line(aes(linetype=Origin), size=1) + scale_fill_brewer(palette = "Paired") + theme_classic(base_size = 17) + theme(legend.position=c(.2, .85), panel.border = element_blank(), axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour = 'black', size=0.5, linetype='solid'), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_x_continuous(breaks=pretty_breaks(n=9)) + labs(linetype="Queen Origin")

```



## In Summary 

* Colony Mass 		= **Higher in Local**   
* Amount of Brood 	= **Higher in Local**   
* Varroa Load 		= **No Difference**   
* Nosema Load		= **Lower in Local**   



## Implications

* Locally-raised queens outperform mass-produced, California queens in their northern environment.   

* This could be evidence for the importance of care in breeding stocks (mass produced vs handmade)   

* **And/Or** This could be evidence for local (genetic) adaptation (imported vs. local)   



## Thank you!

 \begin{picture}(200,200)
 \put(1, 1){\includegraphics[height=30cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture13.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}

## Questions?

 \begin{picture}(200,200)
 \put(1, 1){\includegraphics[height=30cm]
 {/Users/phillipburnham/ComputationalBiology/CompBio/Presentation/Pictures/Picture13.png}}
 \put(1,1){\begin{minipage}[t]{1\linewidth}
 {}
 \end{minipage}}
 \end{picture}
 


